Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-05-09 VegOzone.tex
Refine Section 3.1

\section{Statistical analysis}
\label{sec:stats}
In this section, we assess the climate conditions. To this end, we compute multi-annual mean (referred to as climatology in the following) of environmental key variables (temperature, precipitation, global irradiance) and gound-level \chem{[O_3]} at Svanhovd. For the growing seasons 2018/19, we evaluate the statistical significance of divergances from the norm in these variables (referred to as anomalies). We perform a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at Pallas and Esrange.

@@

\subsubsection{Ozone}
\label{subsubsec:clim_ozone}
In the following, we will derived a ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscadia. To this end, we first calculate a pairwise correlation between the different stations. The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b). Therefore, we select these data to derive a generallized ozone climatology for northern Fennoscandia. The correlation between Svanvik with any of the active monitoring sites in Esrange and Pallas is rather low, 0.4 and 0.6, respectively. Thus, conditions at Svanvik cannot directly be represented by measurements taken at either Esrange or Pallas.

In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, two dimensional density distributions of \chem{[O_3]} for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown. Darker colors indicate higher probability to observe these values. On top of the density distributions, a $10\,\unit{days}$ average of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is displayed together with $1 \sigma$ uncertainties and SE, respectively. Splines have been fitted through the data for easier inter-comparison.

The generalized Fennoscandic \chem{\left<[O_3]\right>} peaks at $48\,\unit{ppb}$ in late April (spring peak). From the northern Fennoscandia climatology, we deduce several features. \chem{\left<[O_3]\right>} declines throughout the growing season until a minimum is reached in late summer/fall ($24-30\,\unit{ppb}$). The turning points lie approximately in June and November. The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum. The pronounced increase in \chem{\left<[O_3]\right>} starts on average one month ahead of the estimated begin of greening ($G_\mathrm{begin}$) (see Section~\ref{subsec:do3se_parameters} for details) and coincidents with rising temperatures and global irradiance. For comparison, we found $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ in units of day of the year (\unit{doy}) for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates to the end of May. This finding is in line with more extensive studies of on-site observations and satellite retrievals \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}. The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is reached in late April ($100-120\,\unit{doy}$). The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$. The minimum occurs at $25\,\unit{ppb}$ in the beginning of September ($240-250\,\unit{doy}$).

The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display a similar pattern as the northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b). The spring peak occurs slightly earlier (at the beginning of April) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology. This is in line with a general decrease in tropospheric ozone with decreasing altitude as Svanvik is located at a lower altitude than the other sites. The earlier decline in \chem{\left<[O_3\right>]} may point to an influence of coniferous vegetation on the removal of ozone since\chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}. In July--September, ozone is occasionally almost completely depleted during nighttime. This may point to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.

\chem{\left<[O_3]\right>} is likely low-biased at Svanvik due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s. To account for this, we derive a bias correction factor by cross-calibrating with the timeseries taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).

>>>>

\section{Statistical analysis}
\label{sec:stats}
In this section, we assess the climate conditions. To this end, we compute multi-annual mean (referred to as climatology in the following) of environmental key variables (temperature, precipitation, global irradiance) and gound-level \chem{[O_3]} at Svanhovd. For the growing seasons 2018/19, we evaluate the statistical significance of divergances from the norm in these variables (referred to as anomalies). We perform a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at Esrange and Pallas.

@@

\subsubsection{Ozone}
\label{subsubsec:clim_ozone}
In the following, we derive an ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscadia. To this end, we first calculate a pairwise correlation between the different stations. The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b). Therefore, we select these data to compute the generallized ozone climatology for northern Fennoscandia. The correlation between Svanvik with any of the active monitoring sites in 2018 (Esrange and Pallas) is rather low, 0.4 and 0.6, respectively. Thus, conditions at Svanvik cannot directly be represented by measurements taken at either Esrange or Pallas.

In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, two dimensional density distributions of \chem{[O_3]} for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown. Darker colors indicate higher probability to observe these values. On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively. Splines have been fitted through the data for easier inter-comparison.

The generalized northern Fennoscandic ozone climatology peaks at $48\,\unit{ppb}$  in late April (day of the year (doy) $100-120$) and declines throughout the growing season until a minimum is reached in late summer/fall ($24-30\,\unit{ppb}$). The turning points lie approximately in June and November. The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum. The most pronounced increase in \chem{\left<[O_3]\right>} coincidents with rising temperatures and global irradiance. The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$. The minimum occurs at $25\,\unit{ppb}$ in the beginning of September ($240-250\,\unit{doy}$).

The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a similar pattern as the generallized northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b). The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology. This is in line with a general decrease in tropospheric ozone with decreasing altitude. The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of greening ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details). In July--September, ozone is occasionally almost completely depleted during nighttime. This hint to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air. The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s. To account for this, we derive a bias correction factor by cross-calibrating with the timeseries taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).
